---
title: "TidyTuesday - Video games dataset"
author: "Pablo Cánovas Tourné"
date: "July 31, 2019"
output:
 html_document: 
  toc: true
  theme: united  
  highlight: tango
editor_options: 
  chunk_output_type: console
---

I recently discovered TidyTuesday, an online data project conducted by
[R for Data Science Online Learning Community](https://github.com/rfordatascience) where a dataset is proposed weekly to people to analyze and find some insights using the tools in the [tidyverse](https://www.tidyverse.org/) R ecosystem to clean, wrangle, tidy and plot a new dataset every Tuesday.

This week the dataset was about video games, and I thought it would be nice to take a look at the data.


```{r setup,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      fig.width = 10, fig.height = 6)
```

```{r}
library(tidyverse)
library(ggthemes)
library(ggridges)
library(ggrepel)
library(viridis)
video_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv")
```


```{r}
df <- video_games %>% 
  mutate(release_date = lubridate::mdy(release_date),
         price = price %>% as.numeric() %>% round()) %>% 
  drop_na(release_date, metascore, price, publisher) %>%
  mutate(owners = owners %>% str_split("[..]") %>%
           map_int(~.x %>% first() %>% str_trim("right") %>% str_replace_all("[,]", "") %>% as.integer())) %>% 
  mutate(publisher = case_when(publisher %>% str_detect("SEGA") ~ "SEGA",
                               publisher %>% tolower() %>% str_detect("square enix") ~ "Square Enix",
                               publisher %>% tolower() %>% str_detect("capcom") ~ "CAPCOM",
                               publisher %>% tolower() %>% str_detect("2k") ~ "2K",
                               TRUE ~ publisher),
         game = case_when(game %>% str_detect("Guacamelee") ~ "Guacamelee!",
                          TRUE ~ game),
         developer = case_when(developer %>% str_detect("Firaxis Games") ~ "Firaxis Games - Aspyr",
                               developer %>% str_detect("Gearbox Software") ~ "Gearbox Software - Aspyr",
                               TRUE ~ developer)) %>% 
  mutate_at(vars(developer, publisher), ~ as.factor(.))
```

Let's start with an overview of the score of the games released by publishers with more titles.


```{r}
big_publishers <- df %>% 
  group_by(publisher) %>% 
  summarise(N_by_publisher = n()) %>% 
  ungroup() %>% 
  top_n(10, N_by_publisher)
df %>%
  filter(publisher %in% big_publishers$publisher) %>% 
  ggplot() + 
  geom_violin(aes(fct_reorder(publisher, metascore, median), metascore,
                   col = fct_reorder(publisher, metascore, median, .desc = TRUE)), 
               alpha = .5, show.legend = F) +
  geom_jitter(aes(publisher, metascore, col = publisher), size = 3, alpha = .4, show.legend = F) +
  coord_flip() + 
  theme(plot.title = element_text(size = 14)) + 
  labs(title = "Score distribution from the 10 publishers with more titles", 
       x = NULL,
       y = "Metascore",
       col = "Publisher")
```

What about the scores of the best rated publishers?

```{r}
top_publishers <- df %>% 
  group_by(publisher) %>% 
  mutate(N_by_publisher = n()) %>% 
  filter(N_by_publisher >= 5) %>% 
  summarise(median_score = median(metascore)) %>% 
  ungroup() %>% 
  top_n(10, median_score)
df %>%
  filter(publisher %in% top_publishers$publisher) %>% 
  ggplot() + 
  geom_violin(aes(fct_reorder(publisher, metascore, median), metascore,
                   col = fct_reorder(publisher, metascore, median, .desc = TRUE)),
               alpha = .5, show.legend = F) +
  geom_jitter(aes(publisher, metascore, col = publisher), size = 3, alpha = .4, show.legend = F) +
  coord_flip() + 
  theme(plot.title = element_text(size = 14), 
        axis.text.y.right = element_text(size = 13)) + 
  labs(title = "Score distribution from top 10 best rated publishers", 
       x = NULL)
```

All of them have a meddian metascore above 80 points!

Now let's take a look at the evolution of the metascores and the amount of titles released through the years

```{r, fig.height = 5}
df %>% 
  mutate(year = release_date %>% lubridate::year() %>% as.factor()) %>% 
  group_by(year) %>% 
  mutate(NGames_year = n()) %>% 
  ggplot() + 
  geom_density_ridges(aes(x = metascore, y = fct_rev(year), fill = NGames_year)) +
  theme(plot.title = element_text(size = 14),
        axis.title.y = element_text(angle = 0, vjust = .5)) + 
  scale_fill_viridis(option = "C") + 
  labs(title = "Metascore evolution through the years", y = NULL, x = "Metascore", fill = "Number of games per year")
```

It seems that with the increasing amount of titles released each year the metascore distribution got more uniform through the years.

It would be interesting to catch a trend in prices too: 

```{r, fig.height = 5}
df %>% 
  mutate(year = release_date %>% lubridate::year() %>% as.factor()) %>% 
  group_by(year) %>% 
  mutate(NGames_year = n()) %>% 
  ggplot() + 
  geom_density_ridges(aes(x = price, y = fct_rev(year), fill = NGames_year)) +
  theme(plot.title = element_text(size = 14),
        axis.title.y = element_text(angle = 0, vjust = .5)) + 
  scale_fill_viridis(option = "C") + 
  labs(title = "Average price evolution through the years",
       y = NULL, x = "Price", fill = "Number of games per year")
```

Effectively, prices seems to have suffered a rise lately.

Another way to look at it, allowing us to quantify *how much* the average price have risen:

```{r, fig.height = 5}
top_publishers <- df %>% 
  group_by(publisher) %>% 
  mutate(N_by_publisher = n()) %>% 
  filter(N_by_publisher >= 5) %>% 
  summarise(median_score = median(metascore)) %>% 
  ungroup() %>% 
  top_n(10, median_score)
df %>% 
  filter(publisher %in% top_publishers$publisher) %>% 
  mutate(release_date = release_date %>% lubridate::ymd(),
         year = release_date %>% lubridate::year() %>% as.factor()) %>% 
  group_by(year, publisher) %>% 
  mutate(mean_price = mean(price)) %>% 
  ggplot() + 
  geom_line(aes(x = year, y = mean_price, col = publisher, group = publisher)) +
  geom_point(aes(x = year, y = mean_price, col = publisher)) + 
  theme(plot.title = element_text(size = 14),
        axis.title.y = element_text(angle = 0, vjust = .5)) + 
  labs(title = "Average price evolution in games from top publishers",
       x = "Year", y = "Average\n price", col = NULL)
```


That rise in price could be justified by an increase in the quality of the products.
So the question is: does the trend in metascores match the trend in prices?

Here we try to adress that question, including also the number of games per year, trying to detect if publishers are now trying to develop *less amount* but *better* video games.


```{r, fig.height = 7}
best_publishers <- df %>% 
  group_by(publisher) %>% 
  mutate(N_by_publisher = n()) %>% 
  summarise(profit = sum(price*owners)/1e6) %>% 
  ungroup() %>% 
  top_n(12, profit)
df %>% 
  filter(publisher %in% best_publishers$publisher) %>%
  mutate(release_date = release_date %>% lubridate::ymd(),
         year = release_date %>% lubridate::year()) %>% 
  group_by(year, publisher) %>% 
  mutate(mean_score = mean(metascore),
         mean_price = mean(price),
         number_of_games = n()) %>% 
  ungroup() %>% 
  select(release_date, publisher, mean_score, mean_price, number_of_games) %>% 
  gather(Var, Value, -release_date, -publisher, -number_of_games) %>% 
  ggplot() + 
  geom_line(aes(x = release_date, y = Value, col = Var)) +
  geom_point(aes(x = release_date, y = Value, size = number_of_games, col = Var)) +
  facet_wrap(~ publisher) + 
  theme(axis.title.y = element_text(angle = 0, vjust = .5), 
        plot.title = element_text(size = 14), 
        plot.subtitle = element_text(size = 12)) +
  labs(title = "Average Score and Price evolution by top publishers", 
       subtitle = "Top publishers based on estimated profits",
       x = "Release date", y = "Score\n & \nPrice") +
  scale_color_tableau()
```

It seems that publishers are raising *prices* but *score* has the opposite trend, especially in companies like 2K.

To finish, let's take a look to best sellers and their metascores:


```{r, fig.height = 7}
df %>% 
  mutate(release_date = release_date %>% lubridate::ymd()) %>% 
  top_n(16, owners) %>%
  arrange(-owners) %>%
  ggplot() + 
  geom_point(aes(x = release_date, y = metascore, col = developer), size = 3, alpha = .8) + 
  geom_text_repel(aes(x = release_date, y = metascore, label = game), point.padding = 0.25) + 
  theme(plot.title = element_text(size = 14), 
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 10),
        axis.title.y = element_text(angle = 0, vjust= 0.5),
        legend.position = "bottom") + 
  labs(x = NULL, y = "Score", title = "Best sellers and their metascore", col = "Developer")
```